---
title: "covid-19 info"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
    flexdashboard::flex_dashboard:
      orientation: rows
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(pacman)
p_load(tidyverse, googlesheets4, myScrapers, RCurl, lubridate, rtweet, patchwork, tmap, leaflet, geojsonio, ggthemes)



```

COVID-19 info and guidance from PHE
===========================================

```{r phe-gov-publications, cache = TRUE}

phe <- get_phe_catalogue(n = 350)

phe %>%
  filter(str_detect(Links, "corona|covid")) %>%
  mutate(Links = str_remove_all(Links, "/government|/guidance"), 
         Links = str_remove_all(Links, "/")) %>%
knitr::kable()

```
---


Data sources
==================================================

### Worldwide data

* World-wide data from JHU datasheets https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data
* These are extracted from the [JHU github repository](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data)
We have combined the data into a single long form table, corrected the dates, calculated country and region totals for cases, deaths and recovered patients (where it exists)
* (Italian data from github repo)
* [Singapore data](https://co.vid19.sg/)


.

---

```{r jhu-data}

cases <- read_csv("https://raw.githubusercontent.com/julianflowers1/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv")
deaths <- read_csv("https://raw.githubusercontent.com/julianflowers1/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv")
recovered <- read_csv("https://raw.githubusercontent.com/julianflowers1/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv")

cases <- cases %>%
  gather(date, value, 5:ncol(.)) %>%
  mutate(date1 = lubridate::mdy(date))

deaths <- deaths %>%
  gather(date, value, 5:ncol(.)) %>%
  mutate(date1 = lubridate::mdy(date))

recovered <- recovered %>%
  gather(date, value, 5:ncol(.)) %>%
  mutate(date1 = lubridate::mdy(date))

combined_data <- bind_cols(cases, deaths = deaths$value, recovered = recovered$value)

combined_data <- combined_data %>%
  select(1:4, date1, cases = value, deaths:recovered) %>%
  gather(metric, value, cases:recovered) %>%
  janitor::clean_names() %>%
  group_by(date1, country_region, metric) %>%
  mutate(total = sum(value)) 

combined_data <- combined_data %>%
  select(country_region, province_state, date1, metric, value, country_total = total, lat, long)

combined_data %>%
  write_csv(paste0("combined-data-", Sys.Date(), ".csv"))

head(combined_data)

```

UK data
============================================

* UK case data from coronavirus tracker
* Tests and death counts from daily twitter feed
* Area based cases from https://www.arcgis.com/home/item.html?id=b684319181f94875a6879bbc833ca3a6

### Cases

```{r}

uk_cases <- readxl::read_excel("DailyConfirmedCases.xlsx") %>%
  add_row()

```

### Tests and deaths

```{r dhsc-data}

source("get_uk_covid_data.R")  

uk_tests <- get_uk_covid_data()

uk_tests

```

### Local area data

```{r la-data}
la_cases <- read_csv("CountyUAs_cases_table.csv")

la_shp <- geojsonio::geojson_read("https://opendata.arcgis.com/datasets/62774ee52e08477f8b65c8b2f15cecb8_0.geojson", what = "sp")

str(la_shp)

eng <- subset(la_shp, substr(ctyua18cd, 1, 1) == "E") ## restrict to England wards
eng@data <- left_join(eng@data, la_cases, by=c("ctyua18cd" = "GSS_CD")) 



```

### Combined UK data

```{r}

combined_uk_data <- uk_cases %>% 
  arrange(., desc(DateVal)) %>%
  bind_cols(filter(uk_tests, date >= "2020-01-31")) %>%
  select(DateVal:deaths) 

combined_uk_data %>%
  write_csv(paste("combined-uk-data-", Sys.Date(), ".csv"))



```

Statistics
==========================

```{r}

options(digits = 2)

combined_uk_stats <- uk_tests %>%
  mutate(daily_cases = pos - lead(pos), 
         daily_tests = total - lead(total), 
         daily_deaths = deaths - lead(deaths), 
         ref_date = date - 1, 
         tests = total, 
         daily_case_rate = round(daily_cases/daily_tests * 1000, 2), 
         daily_death_rate = round(daily_deaths/daily_tests * 1000, 2),
         daily_fatalty_rate = round(daily_deaths/daily_cases * 1000, 2), 
         cumulative_case_rate = round(pos/total * 1000, 2), 
         cumulative_death_rate = round(deaths/total * 1000, 2), 
         cumulative_fatality_rate = round(deaths/pos * 1000, 2)
         )

combined_uk_stats %>%
  knitr::kable()

```

Today's headline numbers
=================================

Row
----------------------------------------
```{r}
today_tests <- combined_uk_stats$daily_tests[1]
today_cases <- combined_uk_stats$daily_cases[1]
today_deaths <- combined_uk_stats$daily_deaths[1]
```

### Cases per day

```{r}
latest_cases <- flexdashboard::valueBox(value = today_cases, caption = "Cases in the last 24 hours", color = "blue", icon = "fa-random")
latest_cases
```


### Tests per day
```{r}
latest_tests <- flexdashboard::valueBox(value = today_tests, caption = "Tests in the last 24 hours", color = "grey", icon = "fa-tag")
latest_tests
```


### Deaths per day

```{r}
latest_deaths <- flexdashboard::valueBox(value = today_deaths, caption = "Deaths in the last 24 hours", color = "red", icon = "fa-tag")
latest_deaths
```



UK plots
=======================================
### Latest UK data
#### Daily trends

```{r, warning=FALSE, message = FALSE}

cases <- combined_uk_stats %>%
  gather(metric, value, c(2:3, total:daily_deaths, daily_case_rate:cumulative_fatality_rate)) %>%
  filter(str_detect(metric, "[Cc]ase"), date >= "2020-02-25") %>%
  ggplot(aes(date, value)) +
  geom_point() +
  geom_line(aes(group = metric)) +
  geom_smooth(se = FALSE, method = "gam", lty = "dotted", lwd = 0.8) +
  facet_wrap(~metric, scales = "free") +
  ggthemes::theme_economist()

```


```{r}

deaths <- combined_uk_stats %>%
  gather(metric, value, c(2:3, total:daily_deaths, daily_case_rate:cumulative_fatality_rate)) %>%
  filter(str_detect(metric, "death"), date >= "2020-02-25") %>%
  ggplot(aes(date, value)) +
  geom_point() +
  geom_line(aes(group = metric)) +
  geom_smooth(se = FALSE, method = "gam", lty = "dotted", lwd = 0.8) +
  facet_wrap(~metric, scales = "free") +
  ggthemes::theme_economist()
```

```{r fig.height=6}

cases / deaths

```

International trends
======================================

#### Trends

```{r fig.height= 12, fig.width= 12}
options(scipen = 999)
combined_data %>%
  ggplot(aes(date1, value, colour = metric)) +
  geom_line(aes(group = country_region)) +
  facet_wrap(~country_region) +
  theme(text = element_text(size = 9)) +
  scale_y_log10() +
  ggthemes::theme_economist_white()


```

International comparisons
=====================================
#### Latest country data

```{r fig.height= 4}

date_range <- range(combined_data$date1)

cases_w <- combined_data %>%
  filter(date1 == date_range[2], metric == "cases") %>%
  group_by(country_region) %>%
  summarise(cum_cases = sum(value)) %>%
  arrange(-cum_cases) %>%
  mutate(date = date_range[2]) %>%
  filter(cum_cases > 29) 


cases_ww <- cases_w %>%
  ggplot(aes(cum_cases, reorder(country_region, cum_cases))) +
  geom_col() +
  labs(y = "") +
  ggthemes::theme_economist_white()

deaths_w <- combined_data %>%
  filter(date1 == date_range[2], metric == "deaths") %>%
  group_by(country_region) %>%
  summarise(cum_deaths = sum(value)) %>%
  arrange(-cum_deaths) %>%
  mutate(date = date_range[2]) %>%
  filter(cum_deaths > 4)


deaths_ww <- deaths_w %>%
  ggplot(aes(cum_deaths, reorder(country_region, cum_deaths))) +
  geom_col() +
  labs(y = "") +
  ggthemes::theme_economist_white()

```

```{r fig.width=8}

cases_ww + deaths_ww

```

LA map
=============================================

```{r}

t <- tm_shape(eng) +
  tm_fill("TotalCases", style = "jenks" , n = 5)  

t %>%
  tmap_leaflet()


```

